---
title       : Data and Models    
---

```{r Setup, include=FALSE, results='hide', warning=FALSE}

library(devtools)
library(ggplot2)

opts_chunk$set(fig.path="assets/wk04-fig-",
               cache.path="cache/wk04-working-",
               dev=c("png","pdf"),
               fig.width=6,
               fig.height=5,
               dpi=300,
               fig.show="hold",
               fig.lp="fig:",
               cache=TRUE,
               par=TRUE,
               echo=TRUE,
               message=FALSE,
               warning=FALSE)

opts_knit$set(kfigr.prefix=TRUE,
              kfigr.link=FALSE)

```

 
# Organs Data

```{r, get-organdata, echo= TRUE}

organs.url <- "https://raw.githubusercontent.com/socviz/soc880/gh-pages/data/organ-procurement.csv"
organs.data <- read.csv(url(organs.url), row.names=1)
dim(organs.data)
head(organs.data)

## For convenience, let R know year is a time measure.
organs.data$year <- as.Date(strptime(organs.data$year, format="%Y"))

```



## Pairs Plot

```{r, warning=FALSE}

install.packages("GGally")
library(GGally)

my.vars <- c("donors", "pop", "gdp",
             "pubhealth", "assault")

#pdf(file="../figures/ggpairs-plot-1.pdf", height=15, width=15)
p <- ggpairs(subset(organs.data, select=my.vars),
        upper=list(continuous="smooth"),
        lower=list(continuous="points"))
p
#dev.off()

```



## Generalized Pairs Plots 

```{r, warning=FALSE}


my.vars <- c("donors", "pop.dens", "gdp",
             "pubhealth", "world", "consent.law",
             "assault", "opt")

# pdf(file="../figures/ggpairs-plot-2.pdf", height=20, width=20)
p <- ggpairs(subset(organs.data, select=my.vars),
        upper=list(continuous="smooth", combo="box"),
        lower=list(continuous="points", combo="dot"))
print(p)
# dev.off()

```


# Looking While Modeling

- Frank Harrell, *Regression Modeling Strategies* (Springer 2001) is an older text, but the visualization principles are good. Two of his useful software libraries for R are `Hmisc` and `rms`.

```{r}

### install.packages("Hmisc")
### install.packages("rms")

library(Hmisc)
library(rms)

vc <- varclus(~ donors + pop.dens + gdp + pubhealth + assault + world + consent.law,
              data=organs.data)
plot(vc)

```



```{r}

plot(naclus(organs.data))

```



```{r}

s <- summary(donors ~ pop.dens + gdp + pubhealth + assault + world + consent.law,
              data=organs.data)

plot(s, cex.labels = 0.6)

```



# Diagnostic Plots for a Poorly-Specified Model

```{r}

out.lm <- lm(donors ~ pop.dens + gdp + pubhealth + assault + consent.law,
             data=organs.data)

summary(out.lm)

```



```{r}

plot(out.lm, which=1)

```



```{r}

plot(out.lm, which=2)

```




```{r}

plot(out.lm, which=3)

```



```{r}

plot(out.lm, which=5)

```



# Plotting results with Hmisc

```{r}

## NB
out.ols <- ols(donors ~ pop.dens + gdp + pubhealth + assault + consent.law,
             data=organs.data)

d <- datadist(organs.data)
options(datadist="d")
options(digits = 2)

plot(summary(out.ols))

```



# Tidy Model Output with `broom`

```{r}

## install.packages("broom")

library(broom)

## more irresponsible model
out.lm <- lm(pubhealth ~ pop.dens + gdp + country,
             data=organs.data)

summary(out.lm)


```



## Model objects have a lot of information in them

```{r}

str(out.lm)

```



## `broom` turns summary model output into data frames

```{r}

out.tidy <- tidy(out.lm, conf.int = TRUE)
class(out.tidy)

head(out.tidy)

```



## It will also "augment" data frames

```{r}

out.aug <- augment(out.lm)
head(out.aug)


```



## Tidy output is easy to quickly plot

```{r}

p <- ggplot(out.tidy, aes(y=estimate,
                          x=reorder(term, estimate)))
p0 <- p + geom_pointrange(aes(ymin=conf.low,
                        ymax=conf.high)) +
    geom_hline(aes(yintercept=0)) +
    coord_flip()

```



## Tidiers are available for many packages

```{r}

library(survival)

out.surv <- survfit(coxph(Surv(time, status) ~ age + sex, data=lung))

out.tidy <- tidy(out.surv)

p <- ggplot(out.tidy, aes(time, estimate))

p + geom_line() +
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)

```



# Model objects have associated `summary` and `plot` methods



## New Zealand Electorate Data

- On GitHub: https://github.com/kjhealy/nz-sex-ratios



# Excursus on Git


## Git is Horrible

- But it can help you keep track of what you did
- It's also where a lot of software development happens
- So it's worth getting a feel for it


# GitHub

## Simplest GitHub 

- You have no account there.
- You download stuff from it.


## Next Simplest GitHub  

- You have an account there.
- You fork (i.e., copy) things on other people's GitHub accounts to
  your GitHub account. 
- Then you clone (i.e., copy, or download) them to your laptop.
- This kind of confusing, doubled-up, garbage terminology is typical of Git. 
 

## Basic One-Person Regular Git Use Case

- You have a file under version control.
- You make a change to it and save the file.
- You stage or `add` a snapshot of the changed files, 
- Then you `commit` the snapshot of changes to the
  repository, with a message saying what you did.
- You `push` that change up to e.g. GitHub.



## Tutorial

- https://try.github.io/levels/1/challenges/1



# Back to the New Zealand Data



