---
title       : Refining Plots    
---

```{r Setup, include=FALSE, results='hide', warning=FALSE}

library(devtools)
library(ggplot2)

opts_chunk$set(fig.path="assets/wk06-fig-",
               cache.path="cache/wk06-working-",
               dev=c("png","pdf"),
               fig.width=6,
               fig.height=5,
               dpi=300,
               fig.show="hold",
               fig.lp="fig:",
               cache=TRUE,
               par=TRUE,
               echo=TRUE,
               message=FALSE,
               warning=FALSE)
			   
opts_knit$set(kfigr.prefix=TRUE,
              kfigr.link=FALSE)
              
knit_hooks$set(plot = function(x, options) {
  paste('<figure class="figure"><img src="',
        opts_knit$get('base.url'), paste(x, collapse = '.'),
        '"><figcaption>', options$fig.cap, '</figcaption></figure>',
        sep = '')
})              

```

# Building up Plots, Again


```{r}

library(ggplot2)
library(scales)
library(MASS)
library(stringr)
library(splines)

theme_set(theme_gray())

```



## ASA Membership & Revenue data

- On Github: [ASA Sections](https://github.com/kjhealy/asa-sections)
- Or manually: 

```{r}

asa.url <- "https://raw.githubusercontent.com/kjhealy/asa-sections/master/data/asa-section-membership.csv"
asa.data <- read.csv((url(asa.url)), header = TRUE)

## If you cloned the github repository, launch R in it and then
## asa.data <-  read.csv("data/asa-section-membership.csv", header=TRUE) 

dim(asa.data)
head(asa.data)


```



## Quick & Dirty Function for custom colors ###

```{r}

my.colors <- function (palette = "cb") {
    cb.palette <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
                    "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
    rcb.palette <- rev(cb.palette)
    bly.palette <- c("#E69F00", "#0072B2", "#999999", "#56B4E9",
                     "#009E73", "#F0E442", "#D55E00", "#CC79A7")
    if (palette == "cb")
        return(cb.palette)
    else if (palette == "rcb")
        return(rcb.palette)
    else if (palette == "bly")
        return(bly.palette)
    else stop("Choose cb, rcb, or bly ony.")
}

```



## Make sure the figures/ directory is available

```{r}

ifelse(!dir.exists(file.path("figures")),
       dir.create(file.path("figures")),
       FALSE)

```



## Starting with the basics again

```{r}

p <- ggplot(asa.data, aes(x=X2014, y=Revenues, label=Sname))

p0 <- p + geom_point()

```



```{r}

p0

```



## Add a smoother

```{r}

p <- ggplot(asa.data, aes(x=X2014, y=Revenues, label=Sname))

p0 <- p + geom_smooth() +
    geom_point()


```


```{r}

p0

```



## Pick out some outliers

```{r}

p <- ggplot(asa.data, aes(x=X2014, y=Revenues, label=Sname))

p0 <- p + geom_smooth() +
    geom_point() +
    geom_text(data=subset(asa.data, Revenues > 7000)) 

```


```{r}

p0

```



## Introduce a third variable ##

```{r}

p <- ggplot(asa.data, aes(x=X2014, y=Revenues, label=Sname))

p0 <- p + geom_smooth() +
    geom_point(aes(color = Journal)) +
    geom_text(data=subset(asa.data, Revenues > 7000)) 

```



```{r}

p0

```




## Change the fitted line ##

```{r}

p <- ggplot(asa.data, aes(x=X2014, y=Revenues, label=Sname))

p0 <- p + geom_smooth(method = "lm",
                  se = FALSE,
                  color = "gray80") +
    geom_point(aes(color = Journal)) +
    geom_text(data=subset(asa.data, Revenues > 7000)) 

```



```{r}

p0

```



## Tidy up the labeled text ##

```{r}

p <- ggplot(asa.data, aes(x=X2014, y=Revenues, label=Sname))

p0 <- p + geom_smooth(method = "lm",
                  se = FALSE,
                  color = "gray80") +
    geom_point(aes(color = Journal)) +
    geom_text(data=subset(asa.data, Revenues > 7000),
              size = 2,
              aes(x=X2014+10,
                  hjust = 0,
                  lineheight = 0.7)) 

```



```{r}

p0

```




## Label the Axes and Scales ##

```{r}

p <- ggplot(asa.data, aes(x=X2014, y=Revenues, label=Sname))

p0 <- p + geom_smooth(method = "lm",
                  se = FALSE,
                  color = "gray80") +
    geom_point(aes(color = Journal)) +
    geom_text(data=subset(asa.data, Revenues > 7000),
              size = 2,
              aes(x=X2014+10,
                  hjust = 0,
                  lineheight = 0.7)) +

   labs(x="Membership",
        y="Revenues",
        color = "Section has own Journal") 

```



```{r}

p0

```



## Fix Tick Marks and Colors ##

```{r}

p <- ggplot(asa.data, aes(x=X2014, y=Revenues, label=Sname))

p0 <- p + geom_smooth(method = "lm",
                  se = FALSE,
                  color = "gray80") +
    geom_point(aes(color = Journal)) +
    geom_text(data=subset(asa.data, Revenues > 7000),
              size = 2,
              aes(x=X2014+10,
                  hjust = 0,
                  lineheight = 0.7)) +
    scale_y_continuous(labels = dollar) +
    scale_color_manual(values = my.colors("bly")) +
    labs(x="Membership",
        y="Revenues",
        color = "Section has own Journal") 

```



```{r}

p0

```



## Add a title and move the legend ##

```{r}

p <- ggplot(asa.data, aes(x=X2014, y=Revenues, label=Sname))

p0 <- p + geom_smooth(method = "lm",
                  se = FALSE,
                  color = "gray80") +
    geom_point(aes(color = Journal)) +
    geom_text(data=subset(asa.data, Revenues > 7000),
              size = 2,
              aes(x=X2014+10,
                  hjust = 0,
                  lineheight = 0.7)) +
    scale_y_continuous(labels = dollar) +
    scale_color_manual(values = my.colors("bly")) +
    labs(x="Membership",
        y="Revenues",
        color = "Section has own Journal") +
    theme(legend.position = "bottom") +
    ggtitle("ASA Sections, Membership vs Revenues")

```



```{r}

p0

```



# Change the theme #

```{r}

theme_set(theme_minimal())

print(p0)


```



## Change the theme again ##

```{r}

theme_set(theme_light())

print(p0)


```



## Moar themes ##

```{r}

library(ggthemes)
theme_set(theme_fivethirtyeight())

print(p0)


```




## Still Moar themes ##

```{r}
theme_set(theme_economist())

p0


```

```{r, echo = FALSE}

theme_set(theme_minimal())

```

# Membership trends over time


```{r}

library(tidyr)
library(dplyr)

yrs <- colnames(asa.data) %in% paste("X", 2005:2015, sep="")

data.m <- subset(asa.data, select = c("Sname", colnames(asa.data)[yrs]))

data.m <- gather(data.m, Year, Members, X2005:X2015)

data.m$Year <- as.integer(str_replace(data.m$Year, "X", ""))

```




# Membership trends over time

## Get the data in shape

```{r}

trend.tab <- data.m %>% group_by(Year) %>%
    mutate(yr.tot = sum(Members, na.rm=TRUE)) %>%
    group_by(Sname) %>%
    na.omit() %>%
    mutate(Ave = mean(Members, na.rm=TRUE),
           Dif = Members - Ave,
           Pct.All = round((Members/yr.tot*100), 2),
           Age = length(Members)) %>%
    group_by(Sname) %>%
    mutate(Index = (Members / first(Members, order_by = Year))*100,
           AveInd = mean(Index))
```


```{r}

index.labs <- trend.tab %>%
    filter(Year == 2015) %>%
    ungroup() %>%
    filter(min_rank(desc(Index)) < 12 | min_rank(desc(Index)) > 44)


index.low <- trend.tab %>%
    filter(Year == 2015) %>%
    ungroup() %>%
    filter(min_rank(Index) < 12)


index.high <- trend.tab %>%
    filter(Year == 2015) %>%
    ungroup() %>%
    filter(min_rank(desc(Index)) < 12)

ind.all <- trend.tab$Sname %in% index.labs$Sname
ind.low <- trend.tab$Sname %in% index.low$Sname
ind.high <- trend.tab$Sname %in% index.high$Sname


trend.tab$Track.all <- ind.all
trend.tab$Track.low <- ind.low
trend.tab$Track.high <- ind.high

```





```{r}

library(quantreg)

p <- ggplot(subset(trend.tab, Age==11 & AveInd < 105),
            aes(x=Year, y=Index, group=Sname, color = Track.low))

p0 <-  p + geom_smooth(method = "rqss", formula = y ~ qss(x), se = FALSE) +
    geom_hline(yintercept = 100) +
    geom_text(data=subset(index.low, Age==11 & AveInd < 105),
              aes(x=Year+0.2, y=Index+rnorm(1, sd=0.8),
                  label=Sname,
                  lineheight=0.8),
              hjust = 0,
              color = "black",
              size = 2.9) +
    expand_limits(x = c(2005:2016)) +
    scale_color_manual(values = my.colors("bly")[c(3, 1)]) +
    scale_x_continuous(breaks = c(seq(2005, 2015, 3))) +
    guides(color = FALSE) +
    ggtitle("Declining Sections. 2005 = 100")

```




```{r}

p0

```



## Simple functions help you out

```{r}

plot.section <- function(section="Culture", x = "Year",
                         y = "Members", data = trend.tab,
                         smooth=FALSE){
    require(ggplot2)
    require(splines)
    ## Note use of aes_string() rather than aes() 
    p <- ggplot(subset(data, Sname==section),
            aes_string(x=x, y=y))

    if(smooth == TRUE) {
        p0 <- p + geom_smooth(color = my.colors("bly")[2],
                              size = 1.2, method = "lm",
                              formula = y ~ ns(x, 3)) +
            scale_x_continuous(breaks = c(seq(2005, 2015, 4))) +
            ggtitle(section)
    } else {
    p0 <- p + geom_line(color=my.colors("bly")[2], size=1.2) +
        scale_x_continuous(breaks = c(seq(2005, 2015, 4))) +
        ggtitle(section)
    }

    print(p0)
}


```



```{r}

plot.section()

```




```{r}

plot.section("Rationality")

```



```{r}

plot.section("OOW")

```



```{r}

plot.section("Sexualities", smooth = TRUE)

```




## Note how this function could be made progressively more general

- E.g. calculate breaks from the data
- Allow `geom_smooth()` arguments to be passed through



```{r}

plot.section2 <- function(section="Culture", x = "Year",
                         y = "Members", data = trend.tab,
                         smooth=FALSE, ...){
    require(ggplot2)
    require(splines)
    ## Note use of aes_string() rather than aes() 
    p <- ggplot(subset(data, Sname==section),
            aes_string(x=x, y=y))

    if(smooth == TRUE) {
        p0 <- p + geom_smooth(color = my.colors("bly")[2],
                              size = 1.2, ...) + 
            scale_x_continuous(breaks = c(seq(2005, 2015, 4))) +
            ggtitle(section)
        } else {
    p0 <- p + geom_line(color=my.colors("bly")[2], size=1.2) +
        scale_x_continuous(breaks = c(seq(2005, 2015, 4))) +
        ggtitle(section)
    }

    print(p0)
}


```



```{r}

plot.section2("Comm/Urban", smooth = TRUE, method = "loess")

```




```{r}

plot.section2("Children", smooth = TRUE, method = "lm", formula = y ~ ns(x, 2))

```



# Another Example: Apple Sales Data

- `git clone https://github.com/kjhealy/apple`

```{r}

apple.url <- "https://raw.githubusercontent.com/kjhealy/apple/master/data/apple-all-products-quarterly-sales.csv"
apple.data <- read.csv((url(apple.url)), header = TRUE)

## If you cloned the github repository, launch R in it and then
## asa.data <-  read.csv("data/asa-section-membership.csv", header=TRUE) 

dim(apple.data)
head(apple.data)


```



```{r}

library(dplyr)
library(ggplot2)
library(tidyr)
library(splines)
library(scales)
library(grid)

## data <- read.csv("data/apple-all-products-quarterly-sales.csv",
## header=TRUE)

apple.data$Date <- seq(as.Date("1998/12/31"), as.Date("2015/12/31"), by = "quarter")

apple.data.m <- gather(apple.data, Product, Sales, iPhone:Mac)

```



```{r}

head(apple.data.m)

```



```{r}

p <- ggplot(subset(apple.data.m, Product!="iPod" & Period>30),
            aes(x=Date, y=Sales, color=Product, fill=Product))
p0 <- p + geom_point(size=1.3) +
    geom_smooth(size=0.8, se=FALSE, method = "loess") +
    theme(legend.position="top") +
    scale_x_date(labels = date_format("%Y"),
                 breaks=date_breaks("year")) +
    scale_colour_manual(values=my.colors()) +
    scale_fill_manual(values=my.colors()) + 
    labs(x="", y="Sales (millions)") 

```



```{r, fig.cap="Quarterly Product Sales"}

p0

```



```{r}

### Convert to time series objects
ipad <- apple.data.m %>%
    group_by(Product) %>%
    filter(Product=="iPad") %>%
    na.omit() %>%
    data.frame(.)

ipad.ts <- ts(ipad$Sales, start=c(2010, 2), frequency = 4)


iphone <- apple.data.m %>%
    group_by(Product) %>%
    filter(Product=="iPhone") %>%
    na.omit() %>%
    data.frame(.)

iphone.ts <- ts(iphone$Sales, start=c(2007, 2), frequency = 4)

mac <- apple.data.m %>%
    group_by(Product) %>%
    filter(Product=="Mac") %>%
    na.omit() %>%
    data.frame(.)

mac.ts <- ts(mac$Sales, start=c(1998, 4), frequency = 4)

```



```{r,  fig.cap="Default stl() decomposition"}

## Loess decomposition 

iphone.stl <- stl(iphone.ts, s.window = "periodic", t.jump = 1)
plot(iphone.stl)

```



## Redraw the STL plot with GGplot

```{r}

iphone.stl2 <- stl(iphone.ts, s.window = 11, t.jump = 1)

ggiphone.stl <- data.frame(iphone.stl2$time.series)

ggiphone.stl$sales <- apple.data$iPhone %>% na.omit()

ind <- is.na(apple.data$iPhone)

ggiphone.stl$Date <- apple.data$Date[!ind]

ggiphone.stl$Product <- "iPhone"


```



```{r}

p <- ggplot(ggiphone.stl, aes(x=Date, y=sales))
p1 <- p + geom_line() + ylab("Data")

p <- ggplot(ggiphone.stl, aes(x=Date, y=trend))
p2 <- p + geom_line() + ylab("Trend")

p <- ggplot(ggiphone.stl, aes(x=Date, y=seasonal))
p3 <- p + geom_line() + ylab("Seasonal")

p <- ggplot(ggiphone.stl, aes(x=Date, y=remainder))
p4 <- p + geom_bar(stat="identity", position="dodge") + ylab("Remainder")

p <- ggplot(ggiphone.stl, aes(x=Date, y=(seasonal/trend)*100))
p5 <- p + geom_line(stat="identity", position="dodge") + ylab("Seasonal/\nTrend (pct)")


```


```{r}

grid.newpage()
vplayout <- function(x, y) viewport(layout.pos.row = x, layout.pos.col = y)

```



```{r, fig.cap="STL decomposition in ggplot"}

pushViewport(viewport(layout = grid.layout(5, 1)))
print(p1, vp = vplayout(1, 1))
print(p2, vp = vplayout(2, 1))
print(p3, vp = vplayout(3, 1))
print(p4, vp = vplayout(4, 1))
print(p5, vp = vplayout(5, 1))

```



# Try it for the Mac and iPad data


